Untitled

### 2022-12-16 draft_report.Rmd
updated details

source(paste0(root,"/haplotype_reconstruction_pipelines.R"))
results_dir <- file.path(root, 'results')

@@

```{r}
#root <- dirname(rstudioapi::getSourceEditorContext()$path)
config <- fread(paste0(root, '/annotations_config.csv'))
print(config)
```

@@

```{r}
bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = T, samples_gen = T)

@@

```{r}
plot_onegeno_test(bxd_results[['ph_geno']], bxd_results[['ph_geno_haploqa']], bxd_results[['map']], ind = 50, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

@@

```{r}
f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = T, samples_gen = F)
```

@@

```{r}
plot_onegeno_test(f2_results[['ph_geno']], f2_results[['ph_geno_haploqa']], f2_results[['map']], ind = 50, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

@@

```{r}
### GigaMUGA mutant
gigamuga_url <- 'http://haploqa-dev.jax.org/tag/Jackson_Lab_Mahaffey_MURGIGV01_20160102_FinalReport.html'
url_domain <- 'http://haploqa-dev.jax.org/'
marker_type <- 'GigaMUGA'
html_table <- read_html(gigamuga_url) %>% html_nodes("a") %>% html_attr("href")
url_list <- paste0(url_domain, html_table[grepl('/sample', html_table)])
summary_df <- sample_summary_scrape(read_html(gigamuga_url), url_list, marker_type)
summary_df$`% No Call` <- as.numeric(gsub("%", "", summary_df$`% No Call`))
summary_df <- summary_df[summary_df$`% No Call` < 10,]

ind_gm <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
#ind_gm <- ind_gm[!ind_gm %in% ind_gm[1:44]] # founder not in UWisc dictionary (SJL)
#ind_gm <- ind_gm[!ind_gm %in% ind_gm[4:24]] # empty - Q2-QP
ind_gm <- ind_gm[!ind_gm %in% c('QU', 'SM', 'SX', 'T5', 'TG', 'TR', 'TS', 'TX', 'UF', 'UR', 'VA', 'VE', 'VG', 'VK', 'VN', 'W4', 'XS', 'Y8', 'Z8', 'ZG')] # screentime error
ind_gm <- ind_gm[!ind_gm %in% c('RY', 'S8', 'SG', 'SR', 'Y2', 'YA', 'YJ', 'YT')] # only one founder
#ind_gm <- ind_gm[!ind_gm %in% c('U2', 'UA', 'UJ' ,'UT', 'W5', 'WD', 'WN', 'XZ', 'Y9', 'YH', 'YS')] # SJL 
#ind_gm <- ind_gm[!ind_gm %in% c('UK', 'UU', 'V4')] #C57BLKS/J
#ind_gm <- ind_gm[!ind_gm %in% c('WX', 'X7', 'X8', 'XF', 'XG', 'XQ', 'XR', 'XY', 'YG', 'YR')] # 129P3
#ind_gm <- ind_gm[!ind_gm %in% c('WY')] # FVB/NJ
#ind_gm <- ind_gm[!ind_gm %in% c('X2', 'X3', 'X4', 'X5', 'XA', 'XB', 'XC', 'XD', 'XJ', 'XK', 'XM', 'XN', 'XT', 'XU', 'XV', 'XW', 'Y2', 'Y3', 'Y4', 'Y5', 'YB', 'YC', 'YD', 'YE', 'YK', 'YM', 'YN', 'YP', 'YU', 'YV', 'YW', 'YX', 'Z5', 'Z6', 'Z7', 'ZC', 'ZD', 'ZE', 'ZF', 'ZM', 'ZN', 'ZP', 'ZQ', 'ZV', 'ZW', 'ZX', 'ZY', 'PY', 'Z3', 'ZB' ,'ZK', 'ZU')] # empty ones

@@

```{r}
### MiniMUGA pipeline
minimuga_url <- 'http://haploqa-dev.jax.org/tag/MiniMUGA.html'
url_domain <- 'http://haploqa-dev.jax.org/'
marker_type <- 'MiniMUGA'
html_table <- read_html(minimuga_url) %>% html_nodes("a") %>% html_attr("href")
url_list <- paste0(url_domain, html_table[grepl('/sample', html_table)])
summary_df <- sample_summary_scrape(read_html(minimuga_url), url_list, marker_type)
ind_mini <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
ind_mini <- ind_mini[ind_mini!='JXX'] # screentime error, skip this one
ind_mini <- ind_mini[ind_mini!='JY9'] # only 1 contributing strain, AIL incompatible, skip this one as well
ind_mini <- ind_mini[ind_mini!='JY6'] # temporary - JXU, JY3, JY5
minimuga_results <- list()
mini_pct_res <- list()
for (ind in ind_mini) {
  #ind <- 'JXN'
  print(ind)
  sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = T, qtl2_file_gen = T)
  df <- ind_geno_comp(sample_mini_res, ind, 'MiniMUGA')

@@

mini_pct_res <- rbindlist(mini_pct_res)
```

@@

```{r}
source(paste0(root,"/app.R"))
shinyApp(ui, server)
```

@@

@@

```{r}
# filepath to the csvs
shiny_pct_fp <- file.path(results_dir, 'shiny_pct_csvs')

# DO
do_pct_df <- fread(file.path(shiny_pct_fp, 'do_truth_comp.csv'))
# CC
cc_pct_df <- fread(file.path(shiny_pct_fp, 'cc_truth_comp.csv'))
# BXD f3
bxd_pct_df <- fread(file.path(shiny_pct_fp, 'bxd_truth_comp.csv'))
# f2
f2_pct_df <- fread(file.path(shiny_pct_fp, 'f2_truth_comp.csv'))
```

@@

The outliers at 80% discordance are samples V6 and VP. Their haplotype reconstruction results appear to be highly heterozygous as CC mice, which explains the high discordance rate. However, it is believed that these two mice samples may have been mistakenly labeled as CC, as the genomes of CC mice are generally homozygous. Further investigation may be needed on the two samples.

@@

```{r}
mean_df <- rbind(t((colMeans(do_pct_df[,2:4]))), t((colMeans(cc_pct_df[,2:4]))), t((colMeans(bxd_pct_df[,2:4]))), t((colMeans(f2_pct_df[,2:4])))) %>% as.data.frame()
colnames(mean_df) <- paste(colnames(mean_df), "mean", sep = "_")

mean_df$sample_type <- c('DO', 'CC', 'F3', 'F2')
mean_df %>% select(sample_type, everything())
colMeans(cc_pct_df[,2:4])
```

@@

@@

```{r}
gm_pct_temp <- rbindlist(gm_pct_res)
gm_plot <- gm_pct_temp %>% group_by(sample_id) %>% summarise(qtl2_pct_diff = sum(haplo_diplotype != qtl2_calls)/n()) %>% as.data.frame()
plot_df <- melt(gm_plot)
plot_df$variable <- 'GigaMUGA Mutant'

ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('GigaMUGA Mutant - percentage of marker difference between genail and haploqa') + xlab('model')

>>>>

source(paste0(root,"/haplotype_reconstruction_pipelines.R"))
results_dir <- file.path(root, 'results')
shiny_pct_fp <- file.path(results_dir, 'shiny_pct_csvs')

@@

```{r}
config <- fread(paste0(root, '/annotations_config.csv'))
print(config)
```

@@

```{r}
bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)

@@

```{r}
plot_onegeno_test(bxd_results[['ph_geno']], bxd_results[['ph_geno_haploqa']], bxd_results[['map']], ind = 5, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

@@

```{r}
f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
```

@@

```{r}
plot_onegeno_test(f2_results[['ph_geno']], f2_results[['ph_geno_haploqa']], f2_results[['map']], ind = 5, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

@@

```{r}
# ### GigaMUGA mutant
gigamuga_url <- 'http://haploqa-dev.jax.org/tag/Jackson_Lab_Mahaffey_MURGIGV01_20160102_FinalReport.html'
url_domain <- 'http://haploqa-dev.jax.org/'
marker_type <- 'GigaMUGA'
html_table <- read_html(gigamuga_url) %>% html_nodes("a") %>% html_attr("href")
url_list <- paste0(url_domain, html_table[grepl('/sample', html_table)])
summary_df <- sample_summary_scrape(read_html(gigamuga_url), url_list, marker_type)
summary_df$`% No Call` <- as.numeric(gsub("%", "", summary_df$`% No Call`))
summary_df <- summary_df[summary_df$`% No Call` < 10,]

ind_gm <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
ind_gm <- ind_gm[!ind_gm %in% c('QU', 'SM', 'SX', 'T5', 'TG', 'TR', 'TS', 'TX', 'UF', 'UR', 'VA', 'VE', 'VG', 'VK', 'VN', 'W4', 'XS', 'Y8', 'Z8', 'ZG')] # screentime error
ind_gm <- ind_gm[!ind_gm %in% c('RY', 'S8', 'SG', 'SR', 'Y2', 'YA', 'YJ', 'YT')] # only one founder

@@

```{r}
mini_res_dir <- file.path(results_dir, 'RDS', 'MiniMUGA_results.rds')
mini_pct_dir <- file.path(shiny_pct_fp, 'mini_pct_comp.csv')

if (!file.exists(mini_res_dir)) {
  ### MiniMUGA pipeline
  minimuga_url <- 'http://haploqa-dev.jax.org/tag/MiniMUGA.html'
  url_domain <- 'http://haploqa-dev.jax.org/'
  marker_type <- 'MiniMUGA'
  html_table <- read_html(minimuga_url) %>% html_nodes("a") %>% html_attr("href")
  url_list <- paste0(url_domain, html_table[grepl('/sample', html_table)])
  summary_df <- sample_summary_scrape(read_html(minimuga_url), url_list, marker_type)
  ind_mini <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
  ind_mini <- ind_mini[ind_mini!='JXX'] # screentime error, skip this one
  ind_mini <- ind_mini[ind_mini!='JY9'] # only 1 contributing strain, AIL incompatible, skip this one as well
  ind_mini <- ind_mini[ind_mini!='JY6'] # temporary - JXU, JY3, JY5
  minimuga_results <- list()
  mini_pct_res <- list()
  for (ind in ind_mini) {
    #ind <- 'JXN'
    print(ind)
    sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = T, qtl2_file_gen = T)
    df <- ind_geno_comp(sample_mini_res, ind, 'MiniMUGA')

@@

  mini_pct_res <- rbindlist(mini_pct_res)
  saveRDS(minimuga_results, mini_res_dir)
} else {
  minimuga_results <- readRDS(mini_res_dir)
}

if (!file.exists(mini_pct_dir)) {
  write.csv(mini_pct_res, mini_pct_dir, row.names = F)
} else {
  mini_pct_res <- fread(mini_pct_dir)
}

@@

```{r}
#source(paste0(root,"/app.R"))
#shinyApp(ui, server)
```

@@

##### GigaMUGA samples:

@@

```{r}
# filepath to the csvs
# DO
do_pct_df <- fread(file.path(shiny_pct_fp, 'do_truth_comp.csv'))
# CC
cc_pct_df <- fread(file.path(shiny_pct_fp, 'cc_truth_comp.csv'))
# BXD f3
bxd_pct_df <- fread(file.path(shiny_pct_fp, 'bxd_truth_comp.csv'))
# f2
f2_pct_df <- fread(file.path(shiny_pct_fp, 'f2_truth_comp.csv'))
```

@@

@@

For the above three plots, there are two significant outliers at 80% discordance showing for CC, which are samples V6 and VP. Their haplotype reconstruction results appear to be highly heterozygous as CC mice, which explains the high discordance rate. However, it is believed that these two mice samples may have been mistakenly labeled as CC, as the genomes of CC mice are generally homozygous. Further investigation may be needed on the two samples.

##### Mutant Mice samples:

@@

The outlier at approximately 70% is sample JXV, which according to its karyotype on HaploQA, this sample has a significant mount of genetic variations for a mutant mice. This could also affect the way qtl2 decides the geno phases and cause the discordance

![Karyotype plot - sample JXV in MiniMUGA](/Users/linc/Desktop/JXV_geno_mini.png)

@@

```{r}
#gm_pct_temp <- rbindlist(gm_pct_res)
#gm_plot <- gm_pct_temp %>% group_by(sample_id) %>% summarise(qtl2_pct_diff = sum(haplo_diplotype != qtl2_calls)/n()) %>% as.data.frame()
#plot_df <- melt(gm_plot)
#plot_df$variable <- 'GigaMUGA Mutant'

#ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('GigaMUGA Mutant - percentage of marker difference between genail and haploqa') + xlab('model')
